### 2019-08-03 paper/diagram.tex
Added original paper diagram

\documentclass{article}

\usepackage[latin1]{inputenc}
\usepackage{tikz}

\usetikzlibrary{shapes,arrows}

\begin{document}
\pagestyle{empty}

% Define block styles
\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
    text width=5em, text centered, rounded corners, minimum height=2em, node distance=2cm]
\tikzstyle{block8} = [rectangle, draw, fill=blue!7, 
text width=5em, text centered, rounded corners, minimum height=2em, node distance=2cm, minimum width=8cm]
\tikzstyle{bigblock10} = [rectangle, draw, fill=red!20, 
	text width=5em, text centered, rounded corners, minimum height=4em, minimum width=10cm]
\tikzstyle{line} = [draw, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
    minimum height=2em]

\begin{tikzpicture}[node distance = 2cm, auto]
	\node [bigblock10] (bert) at (0,0) {bert model};
	\node [block, below of=bert] (SEP1) {'SEP'};
	\node [block, left of=SEP1] (question) {question};
	\node [block, left of=question] (CLS) {'CLS'};
	\node [block, right of=SEP1] (answer) {answer};
	\node [block, right of=answer] (SEP2) {'SEP'};
	\node [block] (classifier) at (-4,2) {classifier token};
	\node [block8] (rest) at (1,2) {unused tokens};
    \path [line] (CLS) -- ([xshift=-4cm]bert.south);
    \path [line] (question) -- ([xshift=-2cm]bert.south);
    \path [line] (SEP1) -- (bert.south);
    \path [line] (answer) -- ([xshift=2cm]bert.south);
    \path [line] (SEP2) -- ([xshift=4cm]bert.south);
    \path [line] ([xshift=-4cm]bert.north) -- (classifier.south);

    \path [line,dashed] ([xshift=-2cm]bert.north) -- ([xshift=-3cm]rest.south);
    \path [line,dashed] ([xshift=-0cm]bert.north) -- ([xshift=-1cm]rest.south);
    \path [line,dashed] ([xshift=2cm]bert.north) -- ([xshift=1cm]rest.south);
    \path [line,dashed] ([xshift=4cm]bert.north) -- ([xshift=3cm]rest.south);

\end{tikzpicture}

\end{document}